TIGLIO. Translations and Images of Greek and Latin Inscriptions Online

### 2016-02-05 source/29/EAGLE2016_camera_ready.tex
paper 29 biblio updated
corrections made to 43, 33, 29

In the fields of epigraphy and archaeology, the areas of digital epigraphy and computational archaeology have benefit
from the use of several of the sensors available in handheld devices. Crowd-sourcing of photographic data and
geo-spatial information, augmented-reality navigation in archaeological spaces and museums, and 3D scanning of
historical artifacts, using smart phones and tablet computers, are few of the exemplar applications of handheld
sensors in epigraphy and archaeology. One common component in all the aforementioned applications is the ability to
record tridimensional data either in the form of geo-spatial coordinates, or in the form of local 3D point coordinates
needed for augmented-reality interaction, or for the construction of triangular meshes of 3D models.

There are several examples in literature that present 3D digitization projects that have been undertaken by museums
including the Epigraphical Museum of Athens \citep{sullivan_analytical_2011, papadaki_accurate_2015}, Museo Arqueológico Nacional de
Madrid \citep{ramirez-sanchez_epigrafidigital:_2014}, Museo Nazionale Romano di Palazzo Altemps \citep{barmpoutis_interactive_2015}, Museo
Geologico Giovanni Capellini di Bologna \citep{abate_valorizzazione_2014}, National Museums Liverpool \citep{cooper_chiswick_2007},
Smithsonian Institution \citep{wachowiak_3d_2009}, and several other museums and institutes \citep{barsanti_3d_2013, landon_petroglyph_2006, levoy_digital_2000}.

Several novel methods for scanning, processing, and analyzing 3D models of inscriptions have been developed, including
methods for text extraction from inscriptions \citep{aswatha_method_2014, sullivan_analytical_2011}, accurate 3D scanning of
inscriptions \citep{papadaki_accurate_2015}, visualization of inscriptions \citep{bozia_open-access_2014}, as well as 3D applications for
other archaeological artifacts \citep{babeu__2011, pollefeys_image-based_2001, malzbender_polynomial_2001, esteban_silhouette_2004}.
Comparative studies of 3D scanning methods for cultural herirage can be found in \citep{pavlidis_methods_2007} and \citep{bohler_3d_2004}.

The aforementioned examples show that the use of 3D technologies in epigraphy and archaeology has been a
well-studied topic over the past two decades. However, there is a notable disconnect between the research on these
technologies and the actual use in the professional epigraphic and archaeological practice as it has been hard for
non-technology-oriented audiences to handle and manipulate tridimensional data, using conventional computer equipment.
Furthermore, without mechanisms for proper user interaction, a 3D model that is projected on a 2D screen is not
significantly advantageous compared to a set of 2D photographs.

@@

\noindent The forms of interactions presented in 1.3.1 and 1.3.2 involved only the motion sensors of a tabled computer without
utilizing the pressure sensors of the screen of the device. The commonly used touch gestures (such as 2-finger twist to
rotate, 2-finger pinch to zoom, and 2-finger translate to move) can be employed in order to enhance the proposed NUI
design. In addition to the aforementioned gestures, a tap gesture could activate regions of interest with additional
modalities of information such as text, images, and metadata. The user can interactively browse the different forms of
data by using intuitive touch gestures as shown in Fig.~\ref{fig:4}. This set of 2D interactions along with the 3D NUI design
presented earlier can compose an intuitive yet powerful workspace for an epigraphist who can now perform digitally
several parts of the epigraphic workflow.

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016cameraready-img006.png}
\caption{Screenshot of our interactive environment. The user interacts with the 3D object, using touch gestures and
selects one of the regions of interests. This action initiates other data tools such as the image viewer or the edge
filter as shown in this example}
\label{fig:4}
\end{figure}

@@

\noindent In this pilot study, a complete set of natural user interactions was designed based on the physical interactions of
epigraphists with real inscriptions. The proposed interactions utilize the existing sensors in a typical tablet
computer or smart phone in order to interactively relight a digitized inscription and manipulate the user's
perspective, using a set of intuitive gestures that imitate the natural interaction with a physical object. In the
proposed design, the epigraphist can ``hold'' a digital inscription, relight it by
reorienting it as a tangible object, observe it from different perspectives, and finally interact with other modalities
by following a set of 2D touch gestures. The prototype system was developed as part of the Digital Epigraphy and
Archaeology (DEA) project using the open-source library VisiNeat for 3D visualization and interaction and is compatible
with iOS, Android, and Microsoft RT tablet and smart phone devices. The interface is available through the web-site of
the project: \url{http://www.digitalepigraphy.org}

In the future we plan to quantitatively evaluate the designed interface by tracking the user activities and analyze
their motion patterns in the 3D space while they are interacting with their handheld device.

>>>>

In the fields of epigraphy and archaeology, the areas of digital epigraphy and computational archaeology have benefited
from the use of several of the sensors available in handheld devices. Crowd-sourcing of photographic data and
geo-spatial information, augmented-reality navigation in archaeological spaces and museums, and 3D scanning of
historical artifacts, using smart phones and tablet computers, are few of the exemplar applications of handheld
sensors in epigraphy and archaeology. One common component in all the aforementioned applications is the ability to
record tridimensional data either in the form of geo-spatial coordinates, or in the form of local 3D point coordinates
needed for augmented-reality interaction, or for the construction of triangular meshes of 3D models.

There are several examples in literature that present 3D digitization projects that have been undertaken by museums
including the Epigraphic Museum of Athens \citep{sullivan_analytical_2011, papadaki_accurate_2015}, Museo Arqueológico Nacional de
Madrid \citep{ramirez-sanchez_epigrafidigital:_2014}, Museo Nazionale Romano di Palazzo Altemps \citep{barmpoutis_interactive_2015}, Museo
Geologico Giovanni Capellini di Bologna \citep{abate_valorizzazione_2014}, National Museums Liverpool \citep{cooper_chiswick_2007},
Smithsonian Institution \citep{wachowiak_3d_2009}, and several other museums and institutes \citep{barsanti_3d_2013, landon_petroglyph_2006, levoy_digital_2000}.

Several novel methods for scanning, processing, and analyzing 3D models of inscriptions have been developed, including
methods for text extraction from inscriptions \citep{aswatha_method_2014, sullivan_analytical_2011}, accurate 3D scanning of
inscriptions \citep{papadaki_accurate_2015}, visualization of inscriptions \citep{bozia_open-access_2014}, as well as 3D applications for
other archaeological artifacts \citep{babeu__2011, pollefeys_image-based_2001, malzbender_polynomial_2001, esteban_silhouette_2004}.
Comparative studies of 3D scanning methods for cultural heritage can be found in \citep{pavlidis_methods_2007} and \citep{bohler_3d_2004}.

The aforementioned examples show that the use of 3D technologies in epigraphy and archaeology has been a
well-studied topic over the past two decades. However, there is a notable disconnect between the research on these
technologies and the actual use in the professional epigraphic and archaeological practice as it has been hard for
non technology-oriented audiences to handle and manipulate tridimensional data, using conventional computer equipment.
Furthermore, without mechanisms for proper user interaction, a 3D model that is projected on a 2D screen is not
significantly advantageous compared to a set of 2D photographs.

@@

\noindent The forms of interactions presented in 1.3.1 and 1.3.2 involved only the motion sensors of a tablet computer without
utilizing the pressure sensors of the screen of the device. The commonly used touch gestures (such as 2-finger twist to
rotate, 2-finger pinch to zoom, and 2-finger translate to move) can be employed in order to enhance the proposed NUI
design. In addition to the aforementioned gestures, a tap gesture could activate regions of interest with additional
modalities of information such as text, images, and metadata. The user can interactively browse the different forms of
data by using intuitive touch gestures as shown in Fig.~\ref{fig:4}. This set of 2D interactions along with the 3D NUI design
presented earlier can compose an intuitive yet powerful workspace for an epigraphist who can now perform digitally
several parts of the epigraphic workflow.

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016cameraready-img006.png}
\caption{Screenshot of our interactive environment. The user interacts with the 3D object, using touch gestures and
selects one of the regions of interests. This action initiates other data tools, such as the image viewer or the edge
filter as shown in this example}
\label{fig:4}
\end{figure}

@@

\noindent In this pilot study, a complete set of natural user interactions was designed based on the physical interactions of
epigraphists with real inscriptions. The proposed interactions utilize the existing sensors in a typical tablet
computer or smart phone in order to interactively relight a digitized inscription and manipulate the user's
perspective, using a set of intuitive gestures that imitate the natural interaction with a physical object. In the
proposed design, the epigraphist can ``hold'' a digital inscription, relight it by
reorienting it as a tangible object, observe it from different perspectives, and finally interact with other modalities
by following a set of 2D touch gestures. The prototype system was developed as part of the Digital Epigraphy and
Archaeology (DEA) project, using the open-source library VisiNeat for 3D visualization and interaction, and is compatible
with iOS, Android, and Microsoft RT tablet and smart phone devices. The interface is available through the web-site of
the project: \url{http://www.digitalepigraphy.org}

In the future, we plan to quantitatively evaluate the designed interface by tracking the user activities and analyze
their motion patterns in the 3D space while they are interacting with their handheld device. 
\nocite{osman_exploratory_2011}

### 2016-02-05 source/33/TIGLIO.tex
corrections made to 43, 33, 29

This comparison confirms also that the usability of resources is measured at a different level when they are made open, and that images and translations have an undeniable higher relevance as an online resource, thus attracting interest also to the transcriptions, while this does not happen the other way around and only those who know what they are looking for will stumble upon an ancient inscription published online. Nothing new in these observations, but this obvious observation is in contrast with the actual situation in which photos are few and translation even fewer. Why so, then? The possible reasons are:

@@

\item people get easily worried by copyright due diligence and find diffucult sometime to trak back who is the author or the copyright owner of a photo

>>>>

This comparison confirms also that the usability of resources is measured at a different level when they are made open, and that images and translations have an undeniable higher relevance as an online resource, thus attracting interest also to the transcriptions, while this does not happen the other way around and only those who know what they are looking for will stumble upon an ancient inscription published online. Nothing new in these observations, but this obvious observation is in contrast with the actual situation in which photos are few and translation even fewer.\footnote{Photographs are very useful  for simple stones inscribed monuments in large lettering (as e.g. Roman monumental texts), but for most longer documentary inscriptions, including most Attic material, after autopsy, squeezes are the most important editorial tool, because photos are mostly 2 D images of 3 D objects.} Why so, then? The possible reasons are:

@@

\item people get easily worried by copyright due diligence and find difficult sometime to trak back who is the author or the copyright owner of a photo
